# Reproducible Research: Peer Assessment 1

## Data
 
* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]

### The variables included in this dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing
    values are coded as `NA`)

* **date**: The date on which the measurement was taken in YYYY-MM-DD
    format

* **interval**: Identifier for the 5-minute interval in which
    measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there
are a total of 17,568 observations in this dataset.

## Loading and preprocessing the data
```{r}
activitydata <- read.csv("activity.csv", colClasses=c("integer", "Date", "integer"))
str(activitydata)
summary(activitydata)
```

## What is mean total number of steps taken per day?
```{r}
steps_by_date <- aggregate(steps ~ date, activitydata, sum)
hist(steps_by_date$steps,breaks=10, col="lightblue", xlab="Steps", labels=TRUE,
     main="Histogram of the total number of steps taken each day")
```

### Mean total number of steps taken per day:
```{r}
mean(steps_by_date$steps)
```

### Median total number of steps taken per day:
```{r}
median(steps_by_date$steps)
```

## What is the average daily activity pattern?

```{r}
steps_by_interval <- aggregate(steps ~ interval, activitydata, mean)
maxsteps_by_interval <- steps_by_interval[which.max(steps_by_interval$steps),]

plot(steps_by_interval, type="l",
     xlab="Time series of the 5-minute interval",
     ylab="The average number of steps taken",
     main="The average daily activity pattern"
    )
i <- maxsteps_by_interval$interval
abline(v=i, col=2, lty=2)
mtext(side=1, text=i, col=2, at=i,line=1)
```

The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:
```{r echo=FALSE}
maxsteps_by_interval
```

## Imputing missing values

The total number of missing values in the dataset:
```{r}
sum(is.na(activitydata))
```


## Are there differences in activity patterns between weekdays and weekends?
