# Reproducible Research: Peer Assessment 1

## Data
 
* Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip) [52K]

### The variables included in this dataset are:

* **steps**: Number of steps taking in a 5-minute interval (missing
    values are coded as `NA`)

* **date**: The date on which the measurement was taken in YYYY-MM-DD
    format

* **interval**: Identifier for the 5-minute interval in which
    measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there
are a total of 17,568 observations in this dataset.

## Loading and preprocessing the data
```{r}
activitydata <- read.csv("activity.csv", colClasses=c("integer", "Date", "integer"))
str(activitydata)
summary(activitydata)
```

## What is mean total number of steps taken per day?
```{r}
steps_bydate <- aggregate(steps ~ date, activitydata, sum)
hist(steps_bydate$steps,breaks=10, col="lightblue", xlab="Steps", labels=TRUE,
     main="Histogram of the total number of steps taken each day")
```

### Mean/Median total number of steps taken per day:
```{r}
mm <- with(steps_bydate, c(mean=mean(steps), median=median(steps)));mm
```

## What is the average daily activity pattern?

```{r}
steps_byinterval <- aggregate(steps ~ interval, activitydata, mean)
maxsteps_byinterval <- steps_byinterval[which.max(steps_byinterval$steps),]

plot(steps_byinterval, type="l",
     xlab="Time series of the 5-minute interval",
     ylab="Average number of steps taken",
     main="Average daily activity pattern"
    )
i <- maxsteps_byinterval$interval
abline(v=i, col=2, lty=2)
mtext(side=1, text=i, col=2, at=i,line=1)
```

The 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:
```{r echo=FALSE}
maxsteps_byinterval
```

## Imputing missing values

The total number of missing values in the dataset:
```{r}
sum(is.na(activitydata))
```
**A new dataset that is equal to the original dataset but with the missing data filled in.(using the mean for that 5-minute interval)**
```{r}
activity <- activitydata
activity$steps <- as.integer(lapply(1:nrow(activity), function(i){
  x <- activity$steps[i]
  if(is.na(x)) with(steps_byinterval, steps[interval == activity$interval[i]])
  else x
}))
```
The total number of missing values in the new dataset:
```{r}
sum(is.na(activity))
```
**With the new dataset,**
### What is mean total number of steps taken per day?
```{r}
steps_bydate_new <- aggregate(steps ~ date, activity, sum)
hist(steps_bydate_new$steps,breaks=10, col="lightblue", xlab="Steps", labels=TRUE,
     main="Histogram of the total number of steps taken each day")
```

### Mean/Median total number of steps taken per day:
```{r}
mm_new <- with(steps_bydate_new, c(mean=mean(steps), median=median(steps)));mm_new
```
This strategy for imputing missing values shifts mean and median lower, using the mean for that 5-minute interval.
```{r}
mm_new - mm
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
Sys.setlocale("LC_TIME", "en_US.UTF-8")
activity$daytype <- as.factor(
  ifelse(weekdays(activity$date,abbreviate=T) %in% c("Sat", "Sun"),
         "weekend", "weekday"))

library(lattice)
with(activity,
  xyplot(steps ~ interval | daytype, 
         aggregate(steps ~ interval + daytype, activitydata, mean), 
         layout = c(1, 2), type = "l",
         xlab = "Interval", ylab = "Number of steps",
         main = "Activity patterns between weekdays and weekends"
         ))
```

The peaks show that the user walks more on weekends.